---
title: "GSS Families"
author: "Joanne Sun, Leqi Sun, Tzu-Ang Su"
date: "October 19, 2020"
abstract: 'In this paper...'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(brms)
library(plyr)
library(broom.mixed)
library(ggmcmc)
library(coda)
library(bayesplot)
library(modelr)
library(tidybayes)
library(gtools)
```

```{r data}
# load the data, store an original copy in the environment and work on a different copy to make sure that we have access to the original data.
orig_data <- read.csv("gss_clean_data_g50.csv")
data <- orig_data

# Divide the age into 6 age groups: under 18, 18-24, 25-34, 35-44, 45-65 and over 65. Notice that for a person whose age is older than the threshold age for each category, the person will be classified into the next category, i.e.: if a person is 24.2 years old, the person belongs to the 25-34 group. 
data <- data %>% mutate(age_group=cut(age, breaks=c(0, 20, 29, 39, 65, Inf), labels=c("under 20", "21-29", "30-39", "40-65", "over 65"))) 

# Change the value of dwelling ownership status to make it easier to display
data$dwelling_own_rent <- mapvalues(data$dwelling_own_rent, from = c("Don't know", "Owned by you or a member of this household, even if it i...", "Rented, even if no cash rent is paid"), to = c("Don't know", "Owned", "Rented"))

# Select the variables of interest in the dataset 
data <- data %>%
  filter(dwelling_own_rent != "Don't know") %>% 
  mutate(dwelling_ownership = dwelling_own_rent,
         gender = as.factor(sex),
         province = as.factor(province),
         highest_education = as.factor(education_highest),
         family_income = as.factor(income_family)) %>% 
  mutate(male = ifelse(gender == "Male", 1, 0)) %>% 
  select(dwelling_ownership, age_group, male, province, highest_education, family_income)

data$dwelling_ownership <- as.numeric(revalue(data$dwelling_ownership, c("Owned" = 1, "Rented" = 0)))

## Take a look at the categorical levels
#levels_income <- levels(data$family_income)
#levels_edu <- levels(data$highest_education)
#levels_age <- levels(data$age_group)
#levels_province <- levels(data$province)

## Plot age group distribution
# page <- ggplot(data = data, mapping = aes(x = age_group)) +
#   theme_light() +
#   geom_bar(position = "dodge", mapping = aes(y = ..prop.., group = 1)) +
#   labs(x = "Age Group", y = "Percentage",
#        title = "Respondents Age Group",
#        subtitle = "The age distribution of the 2017 Family Survey, 2017",
#        caption = "Data: 2017 General Social Survey, Cycle 31: Families, Feburary 2020.")
# page

## Convert categorical variables to numeric for Stan
#data$age_group <- unclass(data$age_group)
#data$highest_education <- unclass(data$highest_education)
#data$family_income <- unclass(data$family_income)
#data$province <- unclass(data$province)
```

```{r postdata, include=FALSE}
# Now create the poststratification data

# Read the 2018 census file downloaded from Census Canada
strat_data <- read_csv("2017_census.csv")
# The original data file has too many age groups, emerge some together, and only keep variables of interest

strat_data <- strat_data %>% 
  mutate(Province = as.factor(COL1),
         Population = COL2,
         Male_under20 = COL3 + COL4,
         Male_29 = COL5 + COL6,
         Male_39 = COL7 + COL8,
         Male_65 = COL9 + COL10 + COL11 + COL12 + COL13,
         Male_over65 = COL14,
         Female_under20 = COL15 + COL16,
         Female_29 = COL17 + COL18,
         Female_39 = COL19 + COL20,
         Female_65 = COL21 + COL22 + COL23 + COL24 + COL25,
         Female_over65 = COL26
         ) %>% 
  select(Province, Population, Male_under20, Male_29, Male_39, Male_65, Male_over65, Female_under20, Female_29, Female_39, Female_65, Female_over65)

# Creat the dataframe for post-stratification, we first extract the variables of interest
age <-  c(rep(c("under 20", "21-29", "30-39", "40-65", "over 65"),length(levels(data$province))*2))

province <-c(rep(levels(data$province), each = 10))

male <- rep(c(rep(1,5),rep(0,5)),10)

count_Alberta <- strat_data %>% 
  filter(Province == "Alberta") %>% 
  select(-Province,
         -Population)
count_BC <- strat_data %>% 
  filter(Province == "British Columbia") %>% 
  select(-Province,
         -Population)
count_Manitoba <- strat_data %>% 
  filter(Province == "Manitoba") %>% 
  select(-Province,
         -Population)
count_NB <- strat_data %>% 
  filter(Province == "New Brunswick") %>% 
  select(-Province,
         -Population)
count_NL <- strat_data %>% 
  filter(Province == "Newfoundland and Labrador") %>% 
  select(-Province,
         -Population)
count_NS <- strat_data %>% 
  filter(Province == "Nova Scotia") %>% 
  select(-Province,
         -Population)
count_ON <- strat_data %>% 
  filter(Province == "Ontario") %>% 
  select(-Province,
         -Population)
count_PEI <- strat_data %>% 
  filter(Province == "Prince Edward Island") %>% 
  select(-Province,
         -Population)
count_QC <- strat_data %>% 
  filter(Province == "Quebec") %>% 
  select(-Province,
         -Population)
count_Saskatchewan <- strat_data %>% 
  filter(Province == "Saskatchewan") %>% 
  select(-Province,
         -Population)

count = c(t(count_Alberta), t(count_BC), t(count_Manitoba), t(count_NB), t(count_NL), t(count_NS), t(count_ON), t(count_PEI), t(count_QC), t(count_Saskatchewan))

# Create the dataframe
poststrat <- as_tibble(cbind(province,age,male, c(count)))

# Convert to appropariate variable types
poststrat <- poststrat %>% 
  mutate(province = as.factor(province),
         age_group = as.factor(age),
         N = as.numeric(V4),
         male = as.numeric(male)) %>% 
  select(province, age_group, male, N)
```



# 1. Introduction

# 2. Data

Data description and where the data came from. The fact it was a opt-in survey....

# 3. Model

We are interested in explaining whether a person owns a house based on age, gender and province of residence. Let $y_{i} = 1$ if the respondent owns a house (or any kinds of dwelling). The model is as the following:

$$Pr(y_i = 1) = \mbox{logit}^{-1}\left(\beta_{0}+\alpha^{age}_{a[i]} + \alpha^{male}_{m[i]} + \alpha^{province}_{p[i]} + \epsilon\right) $$

 $\beta_{0}$ is the global intercept. The notation $a[i]$ refers to the age-group $a$ to which individual $i$ belongs. There are $5$ age groups in total: under 20, 21-29, 30-39, 40-65, and over 65. The notation $m[i]$ refers whether individual $i$ is male. Similarly, $p[i]$ refers to the province individual $i$ resides. The priors for these variables are:
 
$$\beta_0 \sim N(0, 2)$$
$$\alpha^{age}_a \sim N(0, 2)\mbox{ for }a = 1, 2, \ldots, 6$$
The notation $a = 1, 2,\ldots, 6$ represents the age group of each individual.
$$\alpha^{male}_m \sim N(0, 2)\mbox{ for }m = 1, 0$$
$m = 1$ when the individual is male, $m = 0$ when the individual is female.

$$\alpha^{province}_p \sim N(0, 2)\mbox{ for }p = 1, 2, \ldots, 10$$
The notation $p = 1, 2,\ldots, 10$ represents the province of residency for each individual.

$$\epsilon \sim t(3, 0, 2.5) $$
The error term $\epsilon$ has been modeled as following a student t-distribution.

The priors for fixed effects are all $Normal(0,2)$ on $logit$ scale, which means that most of the samples are from about $0.018 - 0.98$ on a natural probability scale.

We chose weakly informative priors for two reasons: first of all, we do not have strong assumptions; secondly, we know that the probability must be greater than $0$ and smaller than $1$. By setting weakly informative priors, we hope that the data could tune the posterior distribution as efficiently as possible.

## Model Checking


```{r model}
options(mc.cores = parallel::detectCores())
## A more Comprehensive model created for pleasure, will not be discussion in this essay
#formula <- dwelling_ownership ~ male + highest_education + (1|age_group) + province + family_income 
#model <- brm(formula = formula, data = data,family = bernoulli(link = "logit"), chains = 4, cores = 4, control = list(adapt_delta=.99), prior = set_prior("normal(0,2)", class = "b"),seed = 1992)

# Create a model
formula1 <- dwelling_ownership ~ male + (1|age_group) + province 
model1 <- brm(formula = formula1, data = data,family = bernoulli(link = "logit"), chains = 4, cores = 4, control = list(adapt_delta=.99), prior = c(prior(normal(0,2), class = Intercept), prior(normal(0,2), class = b)),sample_prior = TRUE, seed = 1992)
prior_summary(model1)
```

Let's do some model checking. Firstly, we do convergence diagnostic to make sure that the MCMC sampling size is big enough. The results of Gelman-Rubin diagnostic shows that the upper confidence limits are either $1$ or very close to $1$, which indicates that the sample chains are converging. It does not flag an issue.

```{r convergencediag}
# Convergence diagnostic, we use Gelman-Rubin Diagnostic, if Upper CI is close to 1, samples converge
model1tranformed <- ggs(model1) 
modelposterior <- as.mcmc(model1) 
gelman.diag(modelposterior)

# Plot the trace plots. The four chains mix well for all of the parameters and therefore, we can conclude no evidence of non-convergence.
mcmc_plot(model1, type = "trace")

mcmc_plot(model1, type = "dens")

#plot(hypothesis(model1, "provinceManitoba = 1"))
```

```{r confusionmatrix}
# The plot shows no evidence of autocorrelation for all model variables in both chains, as the autocorrelation parameters all quickly diminish to around zero.
#mcmc_plot(model1, type = "acf_bar")

# Model checking
Pred_check <- predict(model1, type = "response")
Pred_check <- if_else(Pred_check[,1] > 0.5, 1, 0)
ConfusionMatrix <- table(Pred_check, pull(data, dwelling_ownership)) #`pull` results in a vector
#correct classification rate
ccr<- sum(diag(ConfusionMatrix))/sum(ConfusionMatrix)

# Confusion Matrix 
colnames(ConfusionMatrix) <- c("Predicated: Rent a House", "Predicated: Own a House")
rownames(ConfusionMatrix) <- c("Actual: Rent a House", "Actual: Own a House")
knitr:: kable(ConfusionMatrix)
```

The correct classification rate of the model is `r ccr`. The Confusion Matrix of the model is shown in the table. The model tends to exaggerate the rent rate. 

Before proceed to do post-stratification, one more model check is needed. A posterior predictive check is necessary. In the plot, $y$ represents the observed data, $y_{rep}$ refers to a randomly sample from the posterior distribution. The plot shows that the posterior density is consistent with the observed data.

```{r predictivecheck}
# Set seed
set.seed(1992)

# Set a sample size for prediction
N <- nrow(data)

# Draw a random sample from the full dataset
#data_sub <- data[sample(nrow(data),N),]

# Prediction
pred_model <- posterior_predict(model1, nsamples = 500)
pred_model[is.na(pred_model)] <- mean(pred_model,na.rm = TRUE)


#Posterior predictive check, plot the distribution of observed data vs. 50 simulated datasets. 
samp <- sample(nrow(pred_model), 100)
color_scheme_set("teal")
ppc_dens_overlay(data$dwelling_ownership, pred_model[samp, ])
```

The we do post-stratification based on the age, gender and province. The census data used is from the 2016 Census cycle. The plot shows that the model does a fair job.  
```{r post-stratification}
# Post-stratification, estimation in the population
# Draw a stratified sample

s <- length(levels(poststrat$province)) * length(levels(poststrat$age)) * 2
sample_pop <- sample(1:100, N, prob= (poststrat$N)/strat_data$Population[1], replace=TRUE)
sample <- poststrat[sample_pop,1:3]

# Population prediction
Pop <- posterior_predict(model1, newdata = sample, nsamples = 500)
Pop[is.na(Pop)] <- mean(Pop,na.rm = TRUE)


# Plot the posterior predictive distribution using sampled population datasets vs. observed data
samp <- sample(nrow(Pop), 100)
color_scheme_set("teal")
ppc_dens_overlay(data$dwelling_ownership, Pop[samp, ])

```

# 4. Results


The model estimates and credibel intervals are shown in the table.
```{r result}
# Results
# The estimates is on log scale, use inverse logit function to get the probability estimates
results <- inv.logit(fixef(model1)[,-2])
colnames(results) <- c("Estimate", "2.5%", "97.5%")
rownames(results) <- c("Intercept", "Male", "British Columbia", "Manitoba", "New Brunswick", "Newfoundland and Labrador", "Nova Scotia", "Ontario", "Prince Edward Island", "Quebec", "Saskatchewan")
knitr::kable(results, digits = 3)

```







# 5. Discussions

## 5.1 Weaknesses and Next Steps
An obvious flaw is that no people living in the Territories were sampled in the survey. Indeed there is only less than 0.5% of the Canadian population living in the Territories. However, a comprehensive could provide valuable information about people living in remote regions. The lives in the Territories would be very different from those in the provinces as one can imagine. 

Another drawback is that the gender groups only have the female and the male as options. It can not precisely reflect the demographic reality in the country. Therefore, making population predictions requires great caution. 

The last one is about the analysis. Only gender, age group, and the province of residency were included in the model and post-stratification. Some variables, such as education levels, family income, and marital status, are left out of consideration because we do not have population information. A more comprehensive post-stratification matrix will benefit future studies. Besides, expert knowledge in the relevant field will help identify confounding variables and improve model performance. 



# 6. References

## 6.1 References for the Report

## 6.2 Reference for Data Cleaning

# 7. Appendix
GitHub Link: https://github.com/tomsu0826/g50gssfamiliescycle31
